---
title: "Background substitutions"
author: "Jaeda Patton"
date: "2024-06-18"
output: github_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE, message=FALSE, purl=TRUE}
require("knitr")
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=5, fig.height=5, warning = FALSE, 
                      message = FALSE)
basedir <- file.path("..", "..")

# check for packages and install any that are missing
packages <- c("tidyr", "ggplot2", "Matrix", "stringr", "tibble", "purrr",
              "MASS", "patchwork", "viridis", "doParallel", "ggseqlogo", "ggrepel",
              "forcats", "scales", "dplyr", "onls")
bioconductor_packages <- c()
installed_packages <- packages %in% rownames(installed.packages())
installed_bioc_packages <- bioconductor_packages %in% 
  rownames(installed.packages())

invisible(lapply(paste0('package:', names(sessionInfo()$otherPkgs)), 
                 detach, character.only=TRUE, unload=TRUE))
install.packages(packages[!installed_packages])
if(any(installed_bioc_packages == F)) {
  if(!require("BiocManager", quietly = T)) install.packages("BiocManager")
  BiocManager::install(bioconductor_packages[!installed_bioc_packages])
}

installed_packages <- packages %in% rownames(installed.packages())
installed_bioc_packages <- bioconductor_packages %in% 
  rownames(installed.packages())

# load packages
invisible(lapply(c(bioconductor_packages[installed_bioc_packages], 
                   packages[installed_packages]), 
                 library, character.only=TRUE))


# make output directories
results_dir <- file.path(basedir, "results", "background_substitutions")
if(!dir.exists(results_dir)) dir.create(results_dir)

# plot fontsize
fontsize <- 16
```


This notebook analyzes the effect of the 31 background substitutions on fluorescence, used to produce Figure 7.


### Load data and functions

```{r load, purl = T}
# load general functions
source(file.path(basedir, "scripts", "general_functions.R"))
source(file.path(basedir, "scripts", "background_substitutions_functions.R"))

# reading in complete fluorescence data with functional classification
meanF_data <- read.csv(
  file.path(basedir, "results", "classifying_functional_variants", 
            "meanF_data_fxnal.csv.gz")
)

AncSR1WT_ERE_data <- meanF_data %>% 
  filter(bg == "AncSR1", AA_var == "EGKA", RE == "ERE (GT)")
AncSR2WT_SRE_data <- meanF_data %>% 
  filter(bg == "AncSR2", AA_var == "GSKV", RE == "SRE (AA)")
```


## Analysis

We consider a model where the background substitutions have a nonspecific effect on the affinity of each RH-RE complex. We model a complex's fluorescence as proportional to the fraction of protein bound to DNA. If the affinity of an RH-RE genotype in the AncSR1 background is $K_d$, then its fluorescence $F_1^*$ can be expressed as 

$$
F_1^* = \frac{1}{1 + K_d}
$$

where fluorescence is scaled between 0 and 1. If the background substitutions modify the affinity of a complex by a factor $\alpha$, then fluorescence in the AncSR2 background can be expressed as 

$$
F_2^* = \frac{1}{1 + \alpha K_d}.
$$

Rearranging to express $F_2^*$ in terms of $F_1^*$ gives

$$
F_2^* = \frac{1}{1 + \alpha \left(\frac{1-F_1^*}{F_1^*}\right)}
$$

We will use orthogonal regression to estimate $\alpha$ from the fluorescence data. We will fit the model using only complexes whose fluorescence is significantly greater than that of nonsense variants (active), and whose fluorescence was estimated from the binned sort data. We will use the lower and upper bounds of fluorescence inferred from the RFA models to scale the fluorescence values.

First format and scale the data.

```{r dataformatting, purl = TRUE}
# get lower and upper bounds for each background inferred from RFA models
load(file.path(basedir, "results", "missing_variants", "AncSR1_UL_fit.rda"))
load(file.path(basedir, "results", "missing_variants", "AncSR2_UL_fit.rda"))
AncSR1_UL <- AncSR1_UL_fit$par[(length(AncSR1_UL_fit$par)-1):length(AncSR1_UL_fit$par)]
AncSR2_UL <- AncSR2_UL_fit$par[(length(AncSR2_UL_fit$par)-1):length(AncSR2_UL_fit$par)]
rm(AncSR1_UL_fit, AncSR2_UL_fit)

# Get training data for model;
# only use RH-RE complexes that are active in either background or both
# and which have fluorescence estimates from the binned sort in both backgrounds.
# Also normalize fluorescence to scale between the lower and upper bound of fluorescence,
# inferred from the RFA models

bgepistasismodeldata <-  meanF_data %>%
  select(bg:avg_meanF, type, meanF_REP1:meanF_REP4, sd_meanF, active, functional) %>%
  pivot_wider(names_from = bg, values_from = avg_meanF:functional) %>%
  filter(active_AncSR1 | active_AncSR2) %>%
  filter(type_AncSR1 == "binned", type_AncSR2 == "binned") %>%
  rowwise() %>%
  mutate(se_meanF_AncSR1 = sd_meanF_AncSR1 / sqrt(sum(!is.na(c_across(c(
    meanF_REP1_AncSR1, meanF_REP2_AncSR1, meanF_REP3_AncSR1, meanF_REP4_AncSR1))))),
         se_meanF_AncSR2 = sd_meanF_AncSR2 / sqrt(sum(!is.na(c_across(c(
           meanF_REP1_AncSR2, meanF_REP2_AncSR2, meanF_REP3_AncSR2, meanF_REP4_AncSR2))))),
    pupper_AncSR1 = t.test(c_across(c(
      meanF_REP1_AncSR1, meanF_REP2_AncSR1, meanF_REP3_AncSR1, meanF_REP4_AncSR1)),
      mu = AncSR1_UL[2], alternative = "less", na.action = na.omit)$p.value,
    pupper_AncSR2 = t.test(c_across(c(
      meanF_REP1_AncSR2, meanF_REP2_AncSR2, meanF_REP3_AncSR2, meanF_REP4_AncSR2)),
      mu = AncSR2_UL[2], alternative = "less", na.action = na.omit)$p.value) %>%
  ungroup() %>%
  mutate(F1_norm = (avg_meanF_AncSR1 - AncSR1_UL[1]) / (AncSR1_UL[2] - AncSR1_UL[1]),
         F2_norm = (avg_meanF_AncSR2 - AncSR2_UL[1]) / (AncSR2_UL[2] - AncSR2_UL[1]),
         se1_norm = se_meanF_AncSR1 / (AncSR1_UL[2] - AncSR1_UL[1]),
         se2_norm = se_meanF_AncSR2 / (AncSR2_UL[2] - AncSR2_UL[1]),
         UB_AncSR1 = p.adjust(pupper_AncSR1, method = "fdr") >= 0.1,
         UB_AncSR2 = p.adjust(pupper_AncSR2, method = "fdr") >= 0.1,
         ERE = RE == "ERE (GT)") %>% 
  select(AA_var, RE, F1_norm, F2_norm, se1_norm, se2_norm, ERE, 
         active_AncSR1, active_AncSR2, UB_AncSR1, UB_AncSR2)
```


Fitting the model. Confidence intervals will be obtained by bootstrapping the data and refitting the model.

```{r modelfittingallREs, purl = TRUE, fig.width = 7}
# fit the nonspecific affinity effect model
bgepistasismodel <- onlsmod(F2_norm ~ 1 / (1 + alpha * (1-F1_norm)/F1_norm),
                            bgepistasismodeldata, start = list(alpha = 1), 
                            domain = c(0, 1))

print(bgepistasismodel)
plot(bgepistasismodel, segments = T, fitted.nls = F)

# get r^2 of model
cor(bgepistasismodel$y0, bgepistasismodel$resp)^2

# bootstrapping fits (1000 reps)
if(!file.exists(file.path(results_dir, "allREs_bootstrap.rda"))) {
  # number of variants
  n <- bgepistasismodeldata %>% nrow()
  
  # parallel processing
  cores <- 20
  cl <- parallel::makeCluster(cores, "FORK", outfile = "")
  registerDoParallel(cl)
  
  allREs_bootstrap <- foreach(i = 1:cores, .combine = 'c') %dopar% {
    # data frame to store bootstrapped parameter estimates
    bsvec <- numeric(length = 50)
    
    for(i in 1:50) {
      # sample non-ERE binding variants
      bs <- sample(n, n, replace = T)
      bs_data <- bgepistasismodeldata %>% slice(bs)
      # fit model
      bsvec[i] <- coef(
        onlsmod(F2_norm ~ 1 / (1 + alpha * (1-F1_norm)/F1_norm),
                bs_data, start = list(alpha = 1),
                domain = c(0, 1), verbose = F))
    }
    
    bsvec
  }
  stopCluster(cl)
  save(allREs_bootstrap, 
       file = file.path(results_dir, "allREs_bootstrap.rda"))
} else load(file.path(results_dir, "allREs_bootstrap.rda"))

# get 95% confidence intervals
allREs_95ci <- sort(allREs_bootstrap)[c(25, 976)]

# how many complexes in the dynamic range in AncSR1 are at the upper bound in
# AncSR2, and how many at the lower bound in AncSR1 are in the dynamic range or
# upper bound in AncSR2?
bgepistasismodeldata <- bgepistasismodeldata %>%
  mutate(colors = as.numeric(1*(!active_AncSR1 & active_AncSR2) +
                               2*(active_AncSR1 & !UB_AncSR1 & UB_AncSR2) +
                               3*(UB_AncSR1 & UB_AncSR2)))
bgepistasismodeldata %>%
  filter(active_AncSR1 & !UB_AncSR1) %>%
  count(colors)
bgepistasismodeldata %>%
  filter(!active_AncSR1) %>%
  count(colors)

# plot inferred model
bgepistasismodeldata %>%
  arrange(colors) %>%
  ggplot(aes(x = F1_norm)) +
  geom_point(aes(y = F2_norm, color = factor(colors, levels = 3:0)), size = 2) +
  # 95% confidence intervals
  geom_ribbon(data = data.frame(
    x = seq(0, 1, 0.01),
    lower = bgepistasisfn(seq(0, 1, 0.01), allREs_95ci[1]),
    upper = bgepistasisfn(seq(0, 1, 0.01), allREs_95ci[2])),
    aes(x = x, ymin = lower, ymax = upper),
    fill = "black", alpha = 0.5) +
  scale_color_manual(values = c("3" = "orange", "2" = "magenta", 
                                "1" = "blue", "0" = "gray"),
                     labels = c("UB both", "DR AncSR1, UB AncSR2", 
                                "LB AncSR1, DR/UB AncSR2", "other"),
                     name = "") +
  # plot best fit from data
  geom_function(fun = bgepistasisfn,
                args = list(alpha = coef(bgepistasismodel)),
                color = "black", xlim = c(0, 1), linewidth = 1) +
  # plot functional cutoffs
  geom_vline(xintercept = (AncSR1WT_ERE_data$avg_meanF - AncSR1_UL[1]) / 
               (AncSR1_UL[2] - AncSR1_UL[1]),
             color = "gray40", linetype = "dashed", linewidth = 1)  +
  geom_hline(yintercept = (AncSR2WT_SRE_data$avg_meanF - AncSR2_UL[1]) / 
               (AncSR2_UL[2] - AncSR2_UL[1]),
             color = "gray40", linetype = "dashed", linewidth = 1)  +
  labs(x = "Fluorescence in AncSR1 background",
       y = "Fluorescence in AncSR2 background") +
  xlim(-0.1, 1.1) + ylim(-0.1, 1.1) +
  theme(text = element_text(size = fontsize)) +
  theme_classic()

# fluorescence histograms
bgepistasismodeldata %>%
  mutate(bounds = as.numeric(!active_AncSR1) + as.numeric(2*UB_AncSR1)) %>%
  mutate(bounds = factor(bounds, levels = c(2, 0, 1))) %>%
  ggplot(aes(x = F1_norm, fill = bounds)) +
  geom_histogram(color = "black") +
  scale_fill_manual(values = c("0" = "gray", "1" = "red", "2" = "cyan"),
                    labels = c("DR", "LB", "UB"),
                    name = "") +
  scale_y_continuous(breaks = c(0, 500, 1000)) +
  labs(x = "Fluorescence in AncSR1 background") +
  xlim(-0.1, 1.1) +
  theme(text = element_text(size = fontsize)) +
  theme_classic()

bgepistasismodeldata %>%
  mutate(bounds = as.numeric(!active_AncSR2) + as.numeric(2*UB_AncSR2)) %>%
  mutate(bounds = factor(bounds, levels = c(2, 0, 1))) %>%
  ggplot(aes(x = F2_norm, fill = bounds)) +
  geom_histogram(color = "black") +
  scale_fill_manual(values = c("0" = "gray", "1" = "red", "2" = "cyan"),
                    labels = c("DR", "LB", "UB"),
                    name = "") +
  scale_y_continuous(breaks = c(0, 100, 200)) +
  labs(x = "Fluorescence in AncSR2 background") +
  xlim(-0.1, 1.1) +
  theme(text = element_text(size = fontsize)) +
  theme_classic()
```

The model appears to fit the data well.

What is the relationship between the number of functional RH genotypes per RE in the AncSR1 vs. AncSR2 background?

```{r nvarbg, purl = FALSE}
# count the number of functional variants on each RE
nboundRE <- meanF_data %>%
  group_by(bg, RE) %>%
  summarize(n = sum(functional)) %>%
  ungroup() %>%
  complete(bg, RE, fill = list(n = 0))

# scatter plot
fxnalREscatterplot <- nboundRE %>%
  pivot_wider(names_from = bg, values_from = n) %>%
  ggplot(aes(x = AncSR1, y = AncSR2)) +
  geom_smooth(method = "lm", color = "red") +
  geom_point() +
  geom_text_repel(aes(label = RE), box.padding = 0.15) +
  labs(x = "Number of protein variants bound in AncSR1 background",
       y = "Number of protein variants bound in AncSR2 background") +
  theme(text = element_text(size = fontsize)) +
  theme_classic()
fxnalREscatterplot

# fitting linear model to number of protein variants bound per RE in each bg
nfxnalbglm <- nboundRE %>%
  pivot_wider(names_from = bg, values_from = n) %>%
  column_to_rownames("RE") %>%
  lm(formula = AncSR2 ~ AncSR1, data = .)
summary(nfxnalbglm)
# outlier detection
par(mfrow = c(2,2))
plot(nfxnalbglm)  # ERE is likely an extreme outlier, and SRE a moderate outlier
cooksD <- cooks.distance(nfxnalbglm)
print(cooksD)
outliers <- cooksD[cooksD > (3 * mean(cooksD, na.rm = TRUE))]
print(outliers)
# refit model with outlier removed
nfxnalbglmnooutliers <- nboundRE %>%
  pivot_wider(names_from = bg, values_from = n) %>%
  filter(!RE %in% names(outliers)) %>%
  column_to_rownames("RE") %>%
  lm(formula = AncSR2 ~ AncSR1, data = .)
summary(nfxnalbglmnooutliers)
# prediction interval
nfxnalbgprednooutliers <- predict(
  nfxnalbglmnooutliers, pivot_wider(nboundRE, names_from = bg, values_from = n),
  interval = "prediction")
nfxnalbgprednooutliers <- cbind(
  nfxnalbgprednooutliers, AncSR1 = nboundRE %>% filter(bg == "AncSR1") %>% pull(n))

# replot, show linear fit with outlier removed
nboundRE %>%
  pivot_wider(names_from = bg, values_from = n) %>%
  ggplot(aes(x = AncSR1)) +
  geom_smooth(data = . %>% filter(!RE %in% names(outliers)),
              aes(y = AncSR2),
              method = "lm", color = "red", fullrange = TRUE, se = TRUE) +
  geom_point(aes(y = AncSR2)) +
  geom_text_repel(aes(y = AncSR2, label = RE), box.padding = 0.15) +
  annotate(geom = "text", x = 0, y = 2500, label = paste(
    "adj.~R ^ 2 ==", round(summary(nfxnalbglmnooutliers)$adj.r.squared, 2),
    "~(without~ERE)"),
    parse = TRUE, hjust = 0) +
  annotate(geom = "text", x = 0, y = 2350, label = paste(
    "adj.~R ^ 2 ==", round(summary(nfxnalbglm)$adj.r.squared, 2),
    "~(with~ERE)"),
    parse = TRUE, hjust = 0) +
  labs(x = "Number of protein variants bound\nin AncSR1 background",
       y = "Number of protein variants bound\nin AncSR2 background") +
  theme(text = element_text(size = fontsize)) +
  theme_classic()
```

A linear model fits well for almost all of the REs except for ERE, which is a clear outlier.

Since there is evidence of a specific negative effect of the background substitutions on ERE binding, let's now fit separate background-affinity models for ERE and non-ERE variants. The ratio of $\alpha$ between these two models corresponds to the ERE-specific effect of the background substitutions.

```{r modelfittingEREspecific, purl = TRUE, fig.width = 7}
# fit a model for only non-ERE complexes
bgepistasismodelnonERE <- 
  onlsmod(F2_norm ~ 1 / (1 + alpha * (1-F1_norm)/F1_norm),
          filter(bgepistasismodeldata, !ERE), start = list(alpha = 1),
          domain = c(0, 1))
print(bgepistasismodelnonERE)
plot(bgepistasismodelnonERE, segments = F, fitted.nls = F)
cor(bgepistasismodelnonERE$y0, bgepistasismodelnonERE$resp)^2

# fit a model for only ERE complexes
bgepistasismodelERE <- 
  onlsmod(F2_norm ~ 1 / (1 + alpha * (1-F1_norm)/F1_norm),
          filter(bgepistasismodeldata, ERE), start = list(alpha = 1), 
          domain = c(0, 1))
print(bgepistasismodelERE)
plot(bgepistasismodelERE, segments = F, fitted.nls = F)
cor(bgepistasismodelERE$y0, bgepistasismodelERE$resp)^2

# r^2 for both datasets combined
cor(c(bgepistasismodelnonERE$y0, bgepistasismodelERE$y0),
    c(bgepistasismodelnonERE$resp, bgepistasismodelERE$resp))^2

# bootstrapping fits (1000 reps)
if(!file.exists(file.path(results_dir, "nonERE_ERE_bootstrap.rda"))) {
  # number of ERE/non-ERE binding variants
  n_nonERE <- bgepistasismodeldata %>% filter(!ERE) %>% nrow()
  n_ERE <- bgepistasismodeldata %>% filter(ERE) %>% nrow()
  
  # parallel processing
  cores <- 20
  cl <- parallel::makeCluster(cores, "FORK", outfile = "")
  registerDoParallel(cl)
  
  nonERE_ERE_bootstrap <- foreach(i = 1:cores, .combine = 'rbind') %dopar% {
    # data frame to store bootstrapped parameter estimates
    bsdf <- data.frame(nonEREalpha = rep(0, 50), EREalpha = rep(0, 50))
    
    for(i in 1:50) {
      # sample non-ERE binding variants
      bs_nonERE <- sample(n_nonERE, n_nonERE, replace = T)
      bs_nonERE_data <- bgepistasismodeldata %>% filter(!ERE) %>% slice(bs_nonERE)
      # fit model
      bsdf[i,1] <- coef(
        onlsmod(F2_norm ~ 1 / (1 + alpha * (1-F1_norm)/F1_norm),
                bs_nonERE_data, start = list(alpha = 1),
                domain = c(0, 1), verbose = F))
      
      # sample ERE-binding variants
      bs_ERE <- sample(n_ERE, n_ERE, replace = T)
      bs_ERE_data <- bgepistasismodeldata %>% filter(ERE) %>% slice(bs_ERE)
      # fit model
      bsdf[i,2] <- coef(
        onlsmod(F2_norm ~ 1 / (1 + alpha * (1-F1_norm)/F1_norm),
                bs_ERE_data, start = list(alpha = 1),
                domain = c(0, 1), verbose = F))
    }
    
    bsdf
  }
  stopCluster(cl)
  save(nonERE_ERE_bootstrap, file = file.path(results_dir, "nonERE_ERE_bootstrap.rda"))
} else load(file.path(results_dir, "nonERE_ERE_bootstrap.rda"))

# get 95% confidence intervals
nonERE_95ci <- nonERE_ERE_bootstrap[order(nonERE_ERE_bootstrap[,1]),1][c(25, 976)]
ERE_95ci <- nonERE_ERE_bootstrap[order(nonERE_ERE_bootstrap[,2]),2][c(25, 976)]

# plot inferred models for ERE and non-ERE variants
bgepistasismodeldata %>%
  arrange(ERE) %>%
  ggplot(aes(x = F1_norm)) +
  geom_point(aes(y = F2_norm, color = ERE), size = 2) +
  geom_ribbon(data = data.frame(
    x = seq(0, 1, 0.01),
    lower = bgepistasisfn(seq(0, 1, 0.01), nonERE_95ci[1]),
    upper = bgepistasisfn(seq(0, 1, 0.01), nonERE_95ci[2])),
    aes(x = x, ymin = lower, ymax = upper),
    fill = "black", alpha = 0.3) +
  geom_ribbon(data = data.frame(
    x = seq(0, 1, 0.01),
    lower = bgepistasisfn(seq(0, 1, 0.01), ERE_95ci[1]),
    upper = bgepistasisfn(seq(0, 1, 0.01), ERE_95ci[2])),
    aes(x = x, ymin = lower, ymax = upper),
    fill = "purple2", alpha = 0.3) +
  scale_color_manual(labels = c("FALSE" = "No", "TRUE" = "Yes"), 
                     values = c("gray", "purple2"), name = "ERE") +
  # best fit function non-ERE
  geom_function(fun = bgepistasisfn,
                args = list(alpha = coef(bgepistasismodelnonERE)),
                color = "black", xlim = c(0, 1), linewidth = 1) +
  # best fit function ERE
  geom_function(fun = bgepistasisfn,
                args = list(alpha = coef(bgepistasismodelERE)),
                color = "purple2", xlim = c(0, 1), linewidth = 1) +
  labs(x = "Fluorescence in AncSR1 background",
       y = "Fluorescence in AncSR2 background") +
  xlim(-0.1, 1.1) + ylim(-0.1, 1.1) +
  theme(text = element_text(size = fontsize)) +
  theme_classic()
```

Which ERE variants are significantly different from the model prediction for both the non-ERE-specific and ERE-specific models? Look for variants for which 2*SE of measurement is outside of the 95% confidence intervals of the model.

```{r, purl = FALSE, fig.width = 7}
# Which ERE variants are significantly outside of the 95% confidence interval 
# of the non-RE-specific model or ERE-specific model?
# Look for variants for which 2*SE of fluorescence in both backgrounds is 
# outside of the bounds of the confidence interval.
bgepistasismodeldataERE <- bgepistasismodeldata %>%
  filter(ERE) %>%
  mutate(allREslower95ciF1 = bgepistasisfninv(F2_norm, allREs_95ci[2]),
         allREsupper95ciF1 = bgepistasisfninv(F2_norm, allREs_95ci[1]),
         allREslower95ciF2 = bgepistasisfn(F1_norm, allREs_95ci[2]),
         allREsupper95ciF2 = bgepistasisfn(F1_norm, allREs_95ci[1]),
         allREssiglower = (F1_norm - 2*se1_norm > allREslower95ciF1) & 
           (F2_norm + 2*se2_norm < allREslower95ciF2),
         allREssighigher = (F1_norm + 2*se1_norm < allREsupper95ciF1) & 
           (F2_norm - 2*se2_norm > allREslower95ciF2),
         ERElower95ciF1 = bgepistasisfninv(F2_norm, ERE_95ci[2]),
         EREupper95ciF1 = bgepistasisfninv(F2_norm, ERE_95ci[1]),
         ERElower95ciF2 = bgepistasisfn(F1_norm, ERE_95ci[2]),
         EREupper95ciF2 = bgepistasisfn(F1_norm, ERE_95ci[1]),
         EREsiglower = (F1_norm - 2*se1_norm > ERElower95ciF1) & 
           (F2_norm + 2*se2_norm < ERElower95ciF2),
         EREsighigher = (F1_norm + 2*se1_norm < EREupper95ciF1) & 
           (F2_norm - 2*se2_norm > EREupper95ciF2))

# What fraction of ERE complexes are significantly below the prediction from the
# non-ERE-specific model?
bgepistasismodeldataERE %>% 
  count(allREssiglower) %>%
  mutate(frac = n / sum(n))

# What fraction of ERE complexes are significantly different from the ERE-
# specific model prediction?
bgepistasismodeldataERE %>%
  mutate(sig = EREsiglower | EREsighigher) %>%
  count(sig)

# plot inferred models for ERE and non-ERE variants
bgepistasismodeldata %>%
  left_join(select(bgepistasismodeldataERE, AA_var, RE, EREsiglower, EREsighigher)) %>% 
  replace_na(list(EREsiglower = FALSE, EREsighigher = FALSE)) %>%
  mutate(sig = EREsiglower | EREsighigher) %>%
  arrange(ERE) %>%
  ggplot(aes(x = F1_norm)) +
  geom_point(aes(y = F2_norm, color = ERE, shape = sig), size = 2) +
  # 95% confidence intervals
  geom_ribbon(data = data.frame(
    x = seq(0, 1, 0.01),
    lower = bgepistasisfn(seq(0, 1, 0.01), nonERE_95ci[1]),
    upper = bgepistasisfn(seq(0, 1, 0.01), nonERE_95ci[2])),
    aes(x = x, ymin = lower, ymax = upper),
    fill = "black", alpha = 0.3) +
  geom_ribbon(data = data.frame(
    x = seq(0, 1, 0.01),
    lower = bgepistasisfn(seq(0, 1, 0.01), ERE_95ci[1]),
    upper = bgepistasisfn(seq(0, 1, 0.01), ERE_95ci[2])),
    aes(x = x, ymin = lower, ymax = upper),
    fill = "purple2", alpha = 0.3) +
  scale_color_manual(labels = c("FALSE" = "No", "TRUE" = "Yes"), 
                     values = c("gray", "purple2"), name = "ERE") +
  scale_shape_discrete(labels = c("FALSE" = "No",
                                  "TRUE" = "Yes"),
                       name = "Specific epistasis") +
  geom_function(fun = bgepistasisfn,
                args = list(alpha = coef(bgepistasismodelnonERE)),
                color = "black", xlim = c(0, 1), linewidth = 1) +
  geom_function(fun = bgepistasisfn,
                args = list(alpha = coef(bgepistasismodelERE)),
                color = "purple2", xlim = c(0, 1), linewidth = 1) +
  labs(x = "Fluorescence in AncSR1 background",
       y = "Fluorescence in AncSR2 background") +
  xlim(-0.1, 1.1) + ylim(-0.1, 1.1) +
  theme(text = element_text(size = fontsize)) +
  theme_classic()

# logo plots of ERE-binders with significantly different than predicted 
# fluorescence in AncSR2 from the ERE-specific model
EREepistasislogo <- list("ERE binders with specific\nnegative interactions\nwith background substitutions" = 
                          bgepistasismodeldataERE %>%
                          filter(EREsiglower) %>%
                          pull(AA_var) %>% as.character(),
                        "ERE binders with specific\npositive interactions\nwith background substitutions" = 
                          bgepistasismodeldataERE %>%
                          filter(EREsighigher) %>%
                          pull(AA_var) %>% as.character()) %>%
  ggseqlogo(method = "probability") +
  xlab("RH site") +
  theme(legend.position = "none")
EREepistasislogo
```

